---
title: "Individual deposit growth"
author: "Daniel P. Egan"
date: "12/20/2018"
output: 
  html_document: 
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(scales)
options("getSymbols.warning4.0"=FALSE, 
        "getSymbols.yahoo.warning"=FALSE)
library(quantmod)
library(zoo)
library(lubridate)
library(ggplot2)
library(reshape)
library(dplyr)
library(gridExtra)

thisTicker <- "VTSMX"
initial_deposit <- 100


#' @section Refreshing/re-initalizing the data can take a while, so 
#' I've quarantined that code here.
#' You'll need to set this to 'true' the first time you run it.
if(FALSE){
  source("data_gen.R")  
}

load("data.Rdata")

```


# Setup
What's the relationship between time weighted returns and earnings? Should an investor ever expect to have higher TWR than earnings percent?  

Let's use `r thisTicker` to find out. 

Below we show a steady investor, who always deposits $10 every month into the S&P 500. Sometimes they deposit before a fall, sometimes before a rise. On average, things go up though. 


```{r}
ggplot(growth_data, aes(x=rdate, y=returns_cumul)) + 
  geom_line() + 
  scale_y_continuous(label=dollar, limits = c(0,13))+ 
  ylab("Growth of a dollar") + xlab("") + 
  scale_x_date(date_breaks = "5 year", label = year) + 
  geom_point(data = growth_data %>% filter(!is.na(deposit_date)), 
             aes(x=rdate, y=returns_cumul), col=adjustcolor("orange", alpha.f = 0.3))

```


Over time, those consistent deposits add up. The growth of the market doesn't hurt either. But there are some points where the portfolio is worth less than it's net deposits. 

```{r}
ggplot(veMat, aes(x=rdate)) + 
  geom_line(aes(y = cumulValue), col="blue") + 
  geom_text(data = filter(veMat, rdate==as.Date("2015-01-02")), 
                   aes(x=rdate, 
                       y = cumulValue, 
                       label = "Portfolio Value"), 
            col="blue", vjust = 2, hjust = 1.4) + 
  geom_line(aes(y = netDeposits), col="dark green") +
  geom_text(data = filter(veMat, rdate==as.Date("2015-01-02")), 
                   aes(x=rdate, 
                       y = netDeposits, 
                       label = "Cumulative\ndeposits"), 
            col="dark green", vjust = 1, hjust = -0.5) +
  scale_y_continuous(label=dollar) + 
  ylab("Value") + xlab("") + 
  scale_x_date(date_breaks = "5 year", label = year)

```


What's happening here? Well, let's look at the growth of each individual deposit. The graph below breaks out each individual deposit to show the specific return it generates. 

They all start at $10, but then grow up and down with market movements. Older deposits are red, and younger deposits are yellow. There are a few cases of yellow (newer) deposits being above older (redder) deposits, but they're few and far between. These yellow-above-red are deposits that occurred just before a market bounce. That's luck, not skill thought. We're just doing monthly deposits. 

```{r}


p1 <-
  ggplot(value_streams_long, 
       aes(x=rdate, y=value, group=series, col=series)) + 
  geom_line() + 
  geom_hline(yintercept = initial_deposit) +
  theme(legend.position="none") + 
  scale_color_viridis_d(alpha = 0.3, option = "plasma") + 
  scale_x_date(date_breaks = "5 years", label=year) + 
  scale_y_continuous(label= dollar) + 
  xlab("") +
  ggtitle("By when invested") +
  ylab(paste0("Value of initial ", dollar(initial_deposit))) 
print(p1)
```


This is the same plot, but with all deposits having the same starting point. 

```{r start_aligned}
p2 <-
  ggplot(value_streams_dayaligned, aes(x=days, y=value, group=series, col=series)) + 
  geom_line() + 
  geom_hline(yintercept = initial_deposit) +
  theme(legend.position="none") + 
  scale_color_viridis_d(alpha = 0.3, option = "plasma") + 
  scale_x_continuous(label= function(x) comma(x/365), 
                     breaks = seq(0, 365 * 20, 365 )) + 
  scale_y_continuous(label=dollar) + 
  ggtitle("By how long invested") +
  xlab("Years") +
  ylab(paste0("Value of initial ", dollar(initial_deposit))) 
print(p2)  

```


## The experience
This graph color-codes deposits by whether or not they ever had a loss, and the size of the loss. 

```{r by_min_return_category}


ggplot(value_streams_long, 
       aes(x=rdate, y=value, group=series, col=min_value_cat)) + 
  geom_line() + 
  geom_hline(yintercept = initial_deposit) +
  theme(legend.position="none") + 
  scale_color_viridis_d(alpha = 0.3, option = "D") + 
  scale_x_date(date_breaks = "5 years", label=year) + 
  scale_y_continuous(label= dollar) + 
  xlab("") +
  ggtitle("By when invested") +
  ylab(paste0("Value of initial ", dollar(initial_deposit))) 

```
